
# CodeBook For run_analysis Function.

## Description

The purpose of this function is to clean the data coming from the accelerometers from the Samsung Galaxy S smartphone.

## Usage

run_analysis(folderpath="~")

## Arguments

Argument               | Description
-----------------------|--------------------------------------
folderPath             | The default value for "folderPath" is the default working director, but you can provide it with the required path of the folderthat contains all the required files.

## Steps to reproduce this project
1. Open the R script run_analysis.r using RStudion.
2. Call the function run_analysis() and provide it the path for the folder "UCI HAR Dataset" or you can put the folder in your defualt working directory and it will reach it directly (Note: Default Directory works only with Linux users)
3. Optional for Mac/Windows Users
   + Open the R script run_analysis.r using a text editor.
   + edit the argument for "run_analysis()" function with the folder path of the "UCI HAR Dataset" folder
   
## Description of Data

The features selected for this database come from the accelerometer and gyroscope 3-axial raw signals TimeAcc-XYZ and TimeGyro-XYZ. These time domain signals (prefix 't' to denote time) were captured at a constant rate of 50 Hz. Then they were filtered using a median filter and a 3rd order low pass Butterworth filter with a corner frequency of 20 Hz to remove noise. Similarly, the acceleration signal was then separated into body and gravity acceleration signals (TimeBodyAcc-XYZ and TimeGravityAcc-XYZ) using another low pass Butterworth filter with a corner frequency of 0.3 Hz. 

Subsequently, the body linear acceleration and angular velocity were derived in time to obtain Jerk signals (TimeBodyAccJerk-XYZ and TimeBodyGyroJerk-XYZ). Also the magnitude of these three-dimensional signals were calculated using the Euclidean norm (TimeBodyAccMag, TimeGravityAccMag, TimeBodyAccJerkMag, TimeBodyGyroMag, TimeBodyGyroJerkMag). 

Finally a Fast Fourier Transform (FFT) was applied to some of these signals producing fBodyAcc-XYZ, fBodyAccJerk-XYZ, FrequencyBodyGyro-XYZ, FrequencyBodyAccJerkMag, FrequencyBodyGyroMag, FrequencyBodyGyroJerkMag.

These signals were used to estimate variables of the feature vector for each pattern:  
'-XYZ' is used to denote 3-axial signals in the X, Y and Z directions.

* TimeBodyAcc-XYZ
* TimeGravityAcc-XYZ
* TimeBodyAccJerk-XYZ
* TimeBodyGyro-XYZ
* TimeBodyGyroJerk-XYZ
* TimeBodyAccMag
* TimeGravityAccMag
* TimeBodyAccJerkMag
* TimeBodyGyroMag
* TimeBodyGyroJerkMag
* FrequencyBodyAcc-XYZ
* FrequencyBodyAccJerk-XYZ
* FrequencyBodyGyro-XYZ
* FrequencyBodyAccMag
* FrequencyBodyAccJerkMag
* FrequencyBodyGyroMag
* FrequencyBodyGyroJerkMag


### Variable list and descriptions

Variable name       |           	Description
--------------------|-----------------------------------------
subject 	        | ID the subject who performed the activity for each window sample. Its range is from 1 to 30.
activityName	    | Activity name
activityNumber      | Acitivity Number that refers to an Activity name
Mean()              | Mean value
SD()                | Standard deviation
mad()               | Median absolute deviation 
max()               | Largest value in array
min()               | Smallest value in array
sma()               | Signal magnitude area
energy()            | Energy measure. Sum of the squares divided by the number of values. 
iqr()               | Interquartile range 
entropy()           | Signal entropy
arCoeff()           | Autorregresion coefficients with Burg order equal to 4
correlation()       | correlation coefficient between two signals
maxInds()           | index of the frequency component with largest magnitude
meanFreq()          | Weighted average of the frequency components to obtain a mean frequency
skewness()          | skewness of the frequency domain signal 
kurtosis()          | kurtosis of the frequency domain signal 
bandsEnergy()       |Energy of a frequency interval within the 64 bins of the FFT of each window.
angle()             | Angle between to vectors.

### Dataset structure

```{r} 
str(FinalData)
```

## The transformation process for cleaning the primary data


### reading Data 

* First we started reading data from files:
1. train/X_train
2. test/X_test
3. train/subject_train
4. train/y_train
5. test/subject_test
6. test/y_test
7. features

```{r eval=FALSE} 

X_train <- read.table( paste(folderPath,"/train/X_train.txt", sep = ""))
    X_test <- read.table(paste(folderPath,"/test/X_test.txt", sep = ""))
    Subject_Train <- read.table(paste(folderPath,"/train/subject_train.txt", sep = ""))
    Y_train <- read.table(paste(folderPath,"/train/y_train.txt", sep = ""))
    Subject_Test <- read.table(paste(folderPath,"/test/subject_test.txt", sep = ""))
    Y_test <- read.table(paste(folderPath,"/test/y_test.txt", sep = ""))
    features <- read.table(paste(folderPath,"/features.txt", sep = ""))
    ColumnsNames <- as.array.default(features$V2) #we read in variables names
    ColumnsNames <- as.character(ColumnsNames)

```

### Extracting required columns Names

* At the Coming step we extract the needed columns Names that contains "mean" or "std"

```{r eval=FALSE}

    #### In the Coming steps we extract the needed columns
    #### that contains "mean" or "std"
    counter <- 1
    NeededColumns <- NULL 
    for (x in ColumnsNames) {
        if(grepl("mean",x)  ){
            NeededColumns<- c(NeededColumns,counter)
        }
        counter <- counter + 1
    }
    counter <- 1
    for (x in ColumnsNames) {
        if( grepl("std",x) ){
            NeededColumns<- c(NeededColumns,counter)
        }
        counter <- counter + 1
    }



```

### Subsetting the required columns

* Here we subset only the columns we need with the index we have got from the previous Loob

```{r eval=FALSE}

     X_train<- X_train[,NeededColumns] ##subsetting data
    X_test <- X_test[,NeededColumns]  ##subsetting data
    ColumnsNames<- ColumnsNames[NeededColumns] ##subsetting Columns Names
    merged_train <- cbind(Subject_Train,Y_train,X_train) ##merging the file Subject_Train with X_train with Y_train
    merged_test <- cbind(Subject_Test,Y_test,X_test) ##merging the file Subject_test with X_test with Y_test
    merged_TT <- rbind(merged_train,merged_test) ## here we merge Test data with train data
    ColumnsNames <- append(c("Subject","ActivityNumber"),ColumnsNames) 
    colnames(merged_TT) <- ColumnsNames   ## Naming columns

```

### Referencing Activity Number to it's Name

* Here we started to replace every Activity Number with Activity name putting the new data in Activity

```{r eval=FALSE}

    ActivityName<- NULL
    counter <- 1
    for(x in merged_TT[,2]){
        if(x==1)
            ActivityName <- c(ActivityName,"WALKING")
        else if(x==2)
            ActivityName <- c(ActivityName,"WALKING_UPSTAIRS")
        else if(x==3)
            ActivityName <- c(ActivityName,"WALKING_DOWNSTAIRS")
        else if(x==4)
            ActivityName <- c(ActivityName,"SITTING")
        else if(x==5)
            ActivityName <- c(ActivityName,"STANDING")
        else if(x==6)
            ActivityName <- c(ActivityName,"LAYING")
        else{
            break("label out of order")
        }
        counter <- counter + 1
    }


```

### Getting the mean for every activity and every subject

* Here we subset the second raw of the activity number
and replace it with the activity name that we got
from the previous loob and stored in variable Activity
* At the aggregate function we are getting the mean for every Activity for every subject

```{r eval=FALSE}

    merged_TT <- cbind(merged_TT[1:2],ActivityName,merged_TT[3:81])
    FinalData <- aggregate(merged_TT[4:82],list(Subject = merged_TT$Subject,ActivityName=merged_TT$ActivityName,ActivityNumber=merged_TT$ActivityNumber),mean)
    FinalData <- arrange(FinalData,Subject,ActivityNumber)

```


### Giving variables Descriptive Names

* Here we will give variables decriptive names and at the last step we write the data in .txt file


```{r eval=FALSE}

    names(FinalData)<-gsub("std()", "SD", names(FinalData))
    names(FinalData)<-gsub("mean()", "MEAN", names(FinalData))
    names(FinalData)<-gsub("^t", "Time", names(FinalData))
    names(FinalData)<-gsub("^f", "Frequency", names(FinalData))
    names(FinalData)<-gsub("Acc", "Accelerometer", names(FinalData))
    names(FinalData)<-gsub("Gyro", "Gyroscope", names(FinalData))
    names(FinalData)<-gsub("Mag", "Magnitude", names(FinalData))
    names(FinalData)<-gsub("BodyBody", "Body", names(FinalData))
    write.table(FinalData,"~/run_analysis.txt")

```

## Outputs produced
* Tidy dataset file "TidyDataSamsungGalaxy.txt" file , and will be saved to the default working directory




